---
title: "R Programming"
author: "Apurva Pokhrel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1

Create a vector of size 10, having the values 5,7,9,11,13,13,11,9,7,5. Compute the sum, mean, highest and lowest of these values. Compute the length of this vector? Find the variance and standard deviation for the data of this vector, using the formula for variance and standard deviation. Compare these values by computing the variance and standard deviation using R function. Sort this array values in decreasing order. 

```{r soln1}
# create the vector
vec <- c(5, 7, 9, 11, 13, 13, 11, 9, 7, 5)

# calculate the sum, mean, max and min
vec_sum <- sum(vec)
vec_mean <- mean(vec)
vec_max <- max(vec)
vec_min <- min(vec)

# calculate the length of the vector
vec_len <- length(vec)

# calculate the variance and standard deviation using the formula
vec_var <- sum((vec - mean(vec))^2) / (length(vec) - 1)
vec_sd <- sqrt(vec_var)

# calculate the variance and standard deviation using R functions
vec_var_r <- var(vec)
vec_sd_r <- sd(vec)

# print the results
cat("Variance (formula):", vec_var, "\n")
cat("Standard deviation (formula):", vec_sd, "\n")
cat("Variance (R function):", vec_var_r, "\n")
cat("Standard deviation (R function):", vec_sd_r, "\n")

# sort the vector in decreasing order
vec_sorted <- sort(vec, decreasing = TRUE)

# print the sorted vector
cat("Sorted vector (decreasing order):", vec_sorted, "\n")

```



### Question 2

Create a vector of first 50 even numbers, starting from 2. Also create a vector having values 30 down to 1, as 30, 29, ...,1 

```{r soln2}
# create a vector of the first 50 even numbers starting from 2
vec_even <- seq(from = 2, by = 2, length.out = 50)

# create a vector with values from 30 down to 1
vec_down <- 30:1

# print the vector
cat("The vector of the first 50 even numbers starting from 2 is:", vec_even, "\n")


# print the vector
cat("The vector with values from 30 down to 1 is:", vec_down, "\n")


```

### Question 3
Create a vector of size 10 with 5th and 7th values as missing (store these values as NA). Use the “is.na()” to find locations of missing data. 

```{r soln3}
# Create a vector of size 10 with missing values
my_vec <- c(1, 2, 3, 4, NA, 6, NA, 8, 9, 10)

# Use the "is.na()" function to find the locations of missing data
missing_data <- which(is.na(my_vec))

# Print the locations of missing data
print(missing_data)

```

### Question 4
Create a vector of characters of size 5, consisting of values: “This” “is” “a” “character” “vector”. Find the index of value “is” in the vector using which() or match(). 

```{r soln4}
# Create a vector of characters
my_vec <- c("This", "is", "a", "character", "vector")

# Find the index of "is" in the vector using the "match()" function
index_of_is <- match("is", my_vec)

# Print the index of "is"
print(index_of_is)

```

### Question 5
It is always good to store numerical values rather than textual data. However, while input or output the textual values are easier to understand. An example, for this is as follows in R: 
> Fivepointscale=c(1:5) 
> names(Fivepointscale) = c("Not Satisfactory", "Satisfactory", "Fair", "Good", "Very Good") 
> Feedback = Fivepointscale[c("Good", "Satisfactory")] 
Create a 7-point scale of information input and use this scale to input feedback of 5 students about a question like “Feedback of experience of using an application (Bad, Somewhat bad, not good, ok, good, very good, excellent)”. Find the average of the feedback. 



```{r soln5}
# Create a 7-point scale
Sevenpointscale <- c(1:7)
names(Sevenpointscale) <- c("Bad", "Somewhat bad", "Not good", "OK", "Good", "Very good", "Excellent")

# Input feedback from 5 students
Feedback <- c("Good", "Very good", "Not good", "OK", "Excellent")

# Convert feedback to numerical values based on the 7-point scale
Feedback_numeric <- Sevenpointscale[Feedback]

# Find the average feedback
Avg_feedback <- mean(Feedback_numeric)

# Print the average feedback
print(paste("The average feedback is:", Avg_feedback))
```


### Question 6

Create two strings and concatenate them.

```{r soln6}
# Create two strings
string1 <- "Hello"
string2 <- "world!"

# Concatenate the strings using the paste() function
concatenated_string <- paste(string1, string2)

# Print the concatenated string
print(concatenated_string)

```

### Question 7

Create a long string of words separated by punctuation marks. Replace all the punctuation marks in the string using gsub("[[:punct:]]", "", stringName) function. Find the number of words in the string without punctuation marks. Find the number of distinct words and its count, if possible. 


```{r soln7}
# Create a long string of words separated by punctuation marks
stringName <- "This is a long string, with many words! It also contains several punctuation marks, such as commas, periods, exclamation marks, and question marks?"

# Replace all the punctuation marks in the string using gsub() function
stringName <- gsub("[[:punct:]]", "", stringName)

# Split the string into words using the strsplit() function
words <- strsplit(stringName, " ")[[1]]

# Find the number of words in the string without punctuation marks
num_words <- length(words)

# Print the number of words
print(paste("The number of words in the string without punctuation marks is:", num_words))

# Find the number of distinct words and their count
unique_words <- unique(words)
num_unique_words <- length(unique_words)

# Print the number of distinct words and their count
print(paste("The number of distinct words in the string is:", num_unique_words))
print("The count of each unique word is:")
table(words)

```

### Question 8

Store content in external files for the following types of data in R: 
(i) Vectors (ii) Lists (iii) Arrays (iv) Data frames (v) Factors 
Read those contents into R. Perform operations link sorting on vector data, finding the length of lists and adding data items in list, accessing different elements of array and comparing it to other values, accessing different components of data frames and factors. 

```{r soln8}
print("NOT DONE")
```

### Question 9

Create two matrices of 5*5 size using R, add, subtract and multiply these two matrices. 

```{r soln9}
# Create two matrices
matrix1 <- matrix(1:25, nrow=5)
matrix2 <- matrix(26:50, nrow=5)

# Add the two matrices
add_result <- matrix1 + matrix2

# Subtract the two matrices
sub_result <- matrix1 - matrix2

# Multiply the two matrices
mult_result <- matrix1 %*% matrix2

# Print the results
print("Matrix 1:")
print(matrix1)

print("Matrix 2:")
print(matrix2)

print("Addition result:")
print(add_result)

print("Subtraction result:")
print(sub_result)

print("Multiplication result:")
print(mult_result)

```


### Question 10

Perform transpose of a matrix. 

```{r soln10}
# Create a matrix
mat <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3)

# Print original matrix
cat("Original matrix:\n")
print(mat)

# Transpose matrix
trans_mat <- t(mat)

# Print transposed matrix
cat("Transposed matrix:\n")
print(trans_mat)

```

### Question 11
Find the inverse of a matrix

```{r soln11}
# create a matrix
A <- matrix(c(1, 2, 3, 4), nrow = 2)

# find the inverse of the matrix
A_inv <- solve(A)

# print the inverse of the matrix
A_inv

# verify that A multiplied by its inverse gives the identity matrix
A %*% A_inv

```

### Question 12

Create a list of a factor. Find the occurrences of each factor in the list
```{r soln12}
# Create a list of a factor
my_list <- list(factor("apple"), factor("orange"), factor("banana"), 
                 factor("apple"), factor("banana"), factor("pear"), factor("pear"))

# Find the occurrences of each factor in the list
table(unlist(my_list))
```

### Question 13

Write function to find the largest and smallest values in a 3-dimensional array of size 3*3*3. You should use parameter passing. 

```{r soln13}
find_max_min <- function(arr) {
  max_val <- max(arr)
  min_val <- min(arr)
  return(list(max_val = max_val, min_val = min_val))
}

# Create a 3x3x3 array
arr <- array(c(1:27), dim = c(3, 3, 3))

# Call the function to find the max and min values
result <- find_max_min(arr)

# Print the max and min values
cat("Max value:", result$max_val, "\n")
cat("Min value:", result$min_val, "\n")


```
### Question 14

Find the eigen values and eigen vectors of a symmetric matrix

```{r soln14}
# Create a symmetric matrix
A <- matrix(c(1, 2, 2, 3), nrow = 2)

# Find the eigenvalues and eigenvectors
eigen_values <- eigen(A)$values
eigen_vectors <- eigen(A)$vectors

# Print the results
cat("Eigenvalues:", eigen_values, "\n")
cat("Eigenvectors:\n", eigen_vectors, "\n")

```
### Question 15

Create a table of showing the States of 20 students, assume these students stay in 5 different states. Now create a factor of these states and then compute the frequency of each factor (Hint: You may use factor() and tapply() functions) 
```{r soln15}
# Create a vector of 20 student states
states <- c("West Bengal", "Maharashtra", "Uttar Pradesh", "West Bengal", "Tamil Nadu", "Uttar Pradesh", "West Bengal", "Maharashtra", "Uttar Pradesh", "Tamil Nadu","West Bengal", "Tamil Nadu", "Uttar Pradesh", "Maharashtra", "West Bengal", "Tamil Nadu", "Uttar Pradesh", "Maharashtra", "West Bengal", "Tamil Nadu")

# Convert the states vector to a factor
states_factor <- factor(states)

# Compute the frequency of each state using tapply()
state_freq <- tapply(states_factor, states_factor, length)

# Print the frequency table
state_freq
```
### Question 16

Consider a state wise list of income of few persons. Use factor function to create a frequency division of income into 5 factor classes e.g. 10000-50000; 50000-100000 etc. 

```{r soln16}
# Create a vector of incomes
incomes <- c(55000, 30000, 75000, 90000, 20000, 60000, 40000, 80000, 95000, 35000)

# Use the cut() function to divide the incomes into 5 factor classes
income_factors <- cut(incomes, breaks = c(10000, 50000, 100000, 150000, 200000, 250000), labels = c("10,000-50,000", "50,000-100,000", "100,000-150,000", "150,000-200,000", "200,000-250,000"))

# Compute the frequency of each income factor using table()
income_freq <- table(income_factors)

# Print the frequency table
income_freq


```
### Question 17
Explore different functions in R about strings, arrays, vectors, factors. You may also explore different methods of plotting the data

```{r soln17}
# create some example data
strings <- c("hello", "world", "this is a mixed case string")
array1 <- array(1:12, dim=c(3, 4, 1))
array2 <- array(13:24, dim=c(3, 4, 1))
vector1 <- c(1, 2, 3, 4, 5)
vector2 <- c(6, 7, 8, 9, 10)
factor1 <- factor(c("apple", "banana", "orange", "apple"), levels=c("apple", "banana", "orange"), labels=c("A", "B", "C"))

# concatenate strings
paste(strings, collapse=" ") # "hello world this is a mixed case string"

# convert strings to uppercase or lowercase
toupper(strings) # "HELLO" "WORLD" "THIS IS A MIXED CASE STRING"
tolower(strings) # "hello" "world" "this is a mixed case string"

# extract substrings from strings
substring(strings, 6, 11) # " world" "rld" "s is a"

# get dimensions of arrays
dim(array1) # 3 4 1
dim(array2) # 3 4 1

# apply a function to each element or row/column of an array
apply(array1, c(1,2), sum) # returns a 3x4 matrix of sums of each element in array1

# get length of vectors
length(vector1) # 5
length(vector2) # 5

# get sum of elements in vectors
sum(vector1) # 15
sum(vector2) # 40

# sort vectors in ascending order
sort(vector1) # 1 2 3 4 5
sort(vector2) # 6 7 8 9 10

# get the levels of a factor
levels(factor1) # "apple" "banana" "orange"

# get a frequency table of the levels in a factor
table(factor1) # returns a table of the frequency of each level in factor1

# create scatterplot
plot(vector1, vector2)

# create histogram
hist(vector1)

# create boxplot
boxplot(vector1)

# create bar chart
barplot(table(factor1))

# create a matrix and transpose it
matrix1 <- matrix(1:9, ncol=3)
matrix1_transpose <- t(matrix1)

# calculate the mean of a vector
mean(vector1) # 3

# calculate the median of a vector
median(vector2) # 8

# calculate the variance of a vector
var(vector1) # 2.5

# calculate the standard deviation of a vector
sd(vector2) # 1.581139

# create a pie chart
pie(table(factor1))

# create a line plot
plot(1:5, vector1, type="l")

# create a scatterplot with different colors and shapes for each point
colors <- c("red", "green", "blue", "orange", "purple")
shapes <- c(0, 1, 2, 3, 4)
plot(vector1, vector2, col=colors, pch=shapes)

# create a bar chart with custom colors
colors2 <- c("red", "green", "blue")
barplot(table(factor1), col=colors2)

# create a boxplot with custom labels
boxplot(list(vector1, vector2), names=c("Group 1", "Group2"))

```


